\systemEnding: Building Deep Domain Expertise with Wake/Sleep Bayesian Program Learning

### 2019-05-06 archive.tex
stream of consciousness based on Josh conversation

\title{\systemEnding: A Wake/Sleep Algorithm for Developing Domain Expertise}
\date{}
\begin{document}
\maketitle{}

@@

What would it take to build a machine that can learn everything a
person does over their lifetime?
Although this goal will remain distant for the foreseeable future,
we know that a general learning system like this would need to be able to acquire many different kinds of expertise.
Virtually every child becomes an expert in natural language, motor control, and social interaction, while learning rich
intuitive theories of kinship, taxonomy and physics. Many of those
children will grow into adults with competence in cooking, calculus, tennis, drawing
pictures, or writing software.
Expertise -- the ability to quickly solve new problems in a domain -- is crucial.
Human learning goes far beyond just memorizing a large set of facts or routines,
but involves higher-order skills,
like learning to learn new concepts or
learning to solve new types of problems more quickly and effectively.
Despite great advances in machine learning,
we are still far from
an AI with these abilities.
The modern AI toolkit has given us machines that learn to play challenging games at superhuman levels,
but cannot quickly transfer to similar games like a person would;
or, which can learn to generate convincing English prose,
but which do not then learn to analyze many different languages,
like a field linguist would.
Here, we contribute a computational model that takes a small step toward the goal of building machines that
grow into domain experts,
looking at much simpler kinds of problems that humans can learn to solve and indeed become experts in.

Our model is structured
around the hypothesis that
domain expertise
consists primarily of two ingredients.
First, domain experts have 
explicit, declarative concepts that are powerful yet finely-tuned.
A visual artist has concepts like arcs, spirals, symmetries, and perspectives;
a physicist has concepts like dot products, vector fields, and conservation laws;
and an architect has concepts like arches, supports, and bridges.
Second, experts have implicit, procedural skill in deploying those concepts
to quickly solve new problems:
at a glance, human domain
experts can intuit which compositions of concepts are likely to solve
the task at hand, even before searching for a solution.
Studies of human expertise
find that,
compared to novices,
experts can, at a glance, categorize and classify problems based on the ``deep structure''
of the problem's solution~\cite{chi1981expertise,chi2012seeing}.
In short, we take the stance that expertise means both
having the right explicit concepts, and being able to quickly see how those concepts can be composed into a solution.

Because solutions to many kinds of problems can often be described as
some kind of
program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
our model approaches a specific problem by searching for a program solving
it. It gradually grows its domain-specific knowledge for a
class of problems by assembling a library of code containing concepts
useful for the domain, while jointly training a neural network to quickly
infer, for a specific problem, what code in its library is likely to
solve it. We think of these two learned components as analogous to the
explicit concepts and implicit procedural skill that human experts
develop.  The model is an instance of what in the machine learning
community is called a wake-sleep algorithm, where for us, `waking'
corresponds to solving problems, while a pair of `sleep' cycles
correspond to improving the library of explicit concepts and training
the neural net to improve implicit procedural knowledge of how to
write code.

@@

Our model, called \systemEnding, works through a novel kind
of wake/sleep or `dream' learning~\cite{hinton1995wake}, iterating
through a wake cycle -- where it solves \textbf{tasks} by writing programs (Figure~\ref{threeCycles} top)
-- and a pair of sleep cycles, both of which are loosely biologically
inspired by the different memory consolidation processes that occur during sleep~\cite{rasch2013sleep}. The first sleep cycle, which we refer to as
\textbf{abstraction}, grows the library of code by replying
experiences from waking and consolidating them into new code
abstractions (Figure~\ref{threeCycles} left).  This cycle is loosely inspired by the formation of declarative abstractions
during slow-wave sleep memory consolidation~\cite{DUDAI201520}.  The second
sleep cycle, which we refer to as \textbf{dreaming}, improves the
agents procedural knowledge of how to write code by training a neural network to
help quickly search for programs. The neural net is trained on replayed
experiences as well as `fantasies', or sampled programs, built from the learned
library (Figure~\ref{threeCycles} right).  These two kinds of dreams are inspired by the distinct
episodic replay and hallucination components of dream
sleep~\cite{fosse2003dreaming}.

Viewed as a probabilistic inference problem, \system observes a set of
tasks, written $X$, and infers both a program solving each task, as
well as a prior distribution over programs likely to solve tasks in
the domain (Figure~\ref{threeCycles} middle). This prior is encoded by
a library, written $\mathcal{D}$, which combined with a
learned weight vector $\theta$
defines a generative model over programs
(Appendix~\ref{sampleProgram}).  The neural network learns to invert
this generative model by predicting, conditioned on a task, a
``posterior'' distribution over programs likely to solve that specific
task,
thus functioning as a \textbf{recognition model} and trained jointly with the generative model, in the spirit of the Helmholtz machine~\cite{hinton1995wake}.
Writing $Q(p|x)$ for the approximate
posterior predicted by the recognition model, wake/sleep cycles correspond to iteratively (and
approximately) solving for

\begin{array}{lr}
  \begin{array}{l}  p_x = \argmax_p \probability[x|p]\probability[p|\mathcal{D}^*,\theta^*]\text{, for each task $x\in X$}\end{array}
    &\text{\emph{Wake}}\\\\
\begin{array}{rl}  
\mathcal{D}^* = \argmax_\mathcal{D}&\int\probability\left[\mathcal{D},\phantom{^*}\theta \right]\prod_{x\in X}\sum_{\substack{p\text{ found during waking}}} \probability[x|p]\probability[p|\mathcal{D},\phantom{^*}\theta]\;\text{d}\theta\\
\theta^* = \argmax_\theta&\phantom{\int}\probability\left[\mathcal{D}^*,\theta \right]\prod_{x\in X}\sum_{p\text{ found during waking}} \probability[x|p]\probability[p|\mathcal{D}^*,\theta]
 \end{array}
&\text{\emph{Abstraction (sleep)}}\\\\
\begin{array}{l} Q(p|x)\approx \probability[p|x,\mathcal{D}^*,\theta^*]\end{array}
  &\text{\emph{Dreaming (sleep)}}
\end{array}

\noindent which serves to maximize a lower bound on the posterior over $(\mathcal{D},\theta)$ given $X$ (Appendix~\ref{probabilisticAppendix}).

@@

    \\\bottomrule \\\\
    \multicolumn{2}{c}{
      \begin{tabular}{ccc}
        Dreams before learning&Dreams after learning&Dreams after learning (cherry picked)\\
          \includegraphics[width = 0.3\textwidth]{figures/dreams/beforeLearning25}&
          \includegraphics[width = 0.3\textwidth]{figures/dreams/final25}&
          \includegraphics[width = 0.3\textwidth]{figures/dreams/cherry_picked/montage}
      \end{tabular}
    }
  \end{tabular}
  \caption{\textbf{Top}: 48 (out of 160) LOGO graphics tasks. Agent writes a program controlling a `pen' that draws the target picture. \textbf{Middle}: Example learned library routines. Agent learns parametric routines for drawing families of curves (left) as well as subroutines that take entire programs as input (right). Each row of images on the left is the same code executed with different parameters. Each image on the right is the same code executed with different parameters and with a different subprogram provided as input. \textbf{Bottom}: 25 dreams, or sampled programs, from library before (left column) and after (middle \& right) learning. Blue: where the agent started drawing. Pink: where the agent ended drawing.}\label{logoMaster}
  \end{figure}

@@

  \end{tabular}
  \setlength{\tabcolsep}{6pt} % default is 6pt
  \caption{\textbf{Left:} Learning a language for physical laws starting with recursive list routines like \code{map} and \code{fold}. Agent observes input/outputs of latent functions encoding physical laws (top left, 20 input/outputs per task). Vectors are represented as lists of real numbers. Physical constants expressed in Planck units. \textbf{Right:} Learning a language for recursive list routines starting with only recursion and primitives found in 1959 Lisp. Agent rediscovers \code{map}, \code{fold}, \code{unfold}, etc. For full tasks and learned libraries see Appendix~\ref{appendixMcCarthy}}\label{librariesToLanguages}
  \end{figure}

@@

Our model draws on ideas first elucidated by the
Exploration-Compression family of algorithms~\cite{Dechter:2013:BLV:2540128.2540316,ecc,lazaro2019beyond,DBLP:conf/icml/LiangJK10,solomonoff1989system,schmidhuber2004optimal,ozkural2011teraflop},
which alternate
between searching, or `exploring', a space of solutions, and then
updating their search procedure by compressing solutions found during
exploration.  Both of our sleep phases can be interpreted as a kind of
compression: abstraction aims to compactly refactor code, while the
recognition model aims to encode a program in as few bits as possible,
conditioned on a task. What \system contributes here is the joint capture of
declarative aspects of domain knowledge -- considered in prior EC works -- and
the equally crucial procedural knowledge,
both of which human experts seem to need.

@@

Algorithm~\ref{theWholeThing} specifies how we integrate wake and
sleep cycles.
\begin{algorithm}
  \caption{Full \system algorithm}\label{theWholeThing}
  \begin{algorithmic}[1]
    \State \textbf{function} \systemEnding$(\mathcal{D}, X)$:
    \State {\bfseries Input:} Initial DSL $\mathcal{D}$, tasks $X$
    \State \textbf{Output:} Infinite stream of DSLs, recognition models, and beams
    \State \textbf{Hyperparameters:} Batch size $B$, enumeration timeout $T$, maximum beam size $F$
    \State $\theta\gets $uniform distribution
    \State $\mathcal{F}_x\gets\varnothing $, $\forall x\in X$\Comment{Initialize beams to be empty}
    \While{true}\Comment{Loop over epochs}
    \State shuffle $\gets$ random permutation of $X$\Comment{Randomize minibatches}
    \While{shuffle is not empty}\Comment{Loop over minibatches}
    \State batch $\gets$ first $B$ elements of shuffle\Comment{Next minibatch of tasks}
    \State shuffle $\gets$ shuffle with first $B$ elements removed
    \State $\forall x\in \text{batch}$: $\mathcal{F}_x\gets\mathcal{F}_x\cup \left\{p\;|\;p\in \text{enumerate}(\probability[\cdot |\mathcal{D},\theta],T)\text{ if }\probability[x|p] > 0 \right\}$\Comment{Wake}
    \State Train $Q(\cdot |\cdot )$ to minimize $\mathcal{L}^{\text{MAP}}$ across all $\left\{\mathcal{F}_x \right\}_{x\in X}$\Comment{Dream Sleep}
    \State $\forall x\in \text{batch}$: $\mathcal{F}_x\gets\mathcal{F}_x\cup \left\{p\;|\;p\in \text{enumerate}(Q(\cdot |x),T)\text{ if }\probability[x|p] > 0 \right\}$\Comment{Wake}
    \State $\forall x\in \text{batch}$: $\mathcal{F}_x\gets\text{ top }F\text{ elements of }\mathcal{F}_x\text{ as measured by }\probability[\cdot |x,\mathcal{D},\theta]$\Comment{Keep top $F$ programs}
    \State $\mathcal{D},\theta,\left\{\mathcal{F}_x \right\}_{x\in X}\gets $\textsc{Consolidate}$(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\}_{x\in X})$\Comment{Consolidation Sleep}
    \State \textbf{yield} $(\mathcal{D},\theta)$, $Q$, $\left\{\mathcal{F}_x \right\}_{x\in X}$ \Comment{Yield the updated DSL, recognition model, and solutions found to tasks}
    \EndWhile
    \EndWhile

@@

\begin{algorithm}%[H]
  %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
  \caption{DSL Induction Algorithm}
  \label{grammarInductionAlgorithm}
  \begin{algorithmic}[1]
    \State {\bfseries Input:} Set of beams $\{\mathcal{B}_x\}$
    %         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
    \State \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
    \State $\mathcal{D}\gets$ every primitive in $\{\mathcal{B}_x\}$
    \While{true}         
    \State $\forall p\in \bigcup_{x}\mathcal{B}_x: $ $v_p\gets I\beta(p)$ \Comment{Construct a version space for each program}
    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{B}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(v_p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
    \State Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
    \State Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*(\mathcal{D})) - \|\theta^*(\mathcal{D})\|_0$ \Comment{objective function}
    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{B}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
    \State $\forall x\;:\;\mathcal{B}_x\gets\left\{\text{\textsc{refactor}}(v_p|\mathcal{D})\;:\; p\in \mathcal{B}_x\right\}$\Comment{Refactor beams in terms of new DSL}
    \EndWhile
  \end{algorithmic}
\end{algorithm}

 \subsubsection{Refactoring code with version spaces}\label{appendixVersion}
 Formally, a version space is either:
\begin{itemize}
\item A deBuijn\footnote{deBuijn indices are an alternative way of naming variables in $\lambda$-calculus. When using deBuijn indices, $\lambda$-abstractions are written \emph{without} a variable name, and variables are written as the count of the number of $\lambda$-abstractions up in the syntax tree the variable is bound to. For example, $\lambda x.\lambda y. (x\;y)$ is written $\lambda\lambda (\$1\;\$0)$ using deBuijn indices. See~\cite{pierce} for more details.} index: written $\$i$, where $i$ is a natural number
\item An abstraction: written $\lambda v$, where $v$ is a version space
\item An  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
\item A  union: $\uplus V$, where $V$ is a set of version spaces
\item The empty set, $\varnothing$
\item The set of all $\lambda$-calculus expressions, $\Lambda$
\end{itemize}
The purpose of a version space to compactly represent a set of programs.
We refer to this set as the \textbf{extension} of the version space:
\begin{definition}
  The \textbf{extension} of a version space $v$ is written $\denotation{v}$
  and is defined recursively as:
  \begin{align*}
    \denotation{\$i}& = \left\{\$i \right\}&
    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}&
    \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}&
    \denotation{\varnothing}& = \varnothing&
    \denotation{\Lambda}& = \Lambda
    \end{align*}
\end{definition}
Version spaces also support efficient membership checking, which we
write as  $e\in \denotation{v}$.
Important for our purposes, it is also efficient to refactor the
members of a version space's extension in terms of a new DSL.
We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
\begin{align*}
  \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
    e\text{, if $e\in \mathcal{D}$ and $e\in \denotation{v}$. Exploits the fact that $\denotation{e}\in v$ can be efficiently computed.}\\
    \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
  \end{cases}
\end{align*}\vspace{-\baselineskip}%
\begin{align*}
  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}&%\\\nonumber
  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\\nonumber
  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})&%\\\nonumber
  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\text{size}(e|\mathcal{D})\nonumber
  \end{align*}
where $\text{size}(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
is the size of the syntax tree of $e$,
when members of $\mathcal{D}$ are counted as having size 1.
Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\text{size}(p|\mathcal{D})$.

@@

\subsubsection{Computational complexity of DSL learning}

>>>>

\title{\systemEnding: Building Deep Domain Expertise with Wake/Sleep Bayesian Program Learning}
\date{}
\begin{document}
\maketitle{}

\vspace{-1cm}

@@

What would it take to build a machine that can learn everything a
person does over their lifetime?
Although this goal will remain distant for the foreseeable future,
we know that a general learning system like this would need to be able to acquire many different kinds of expertise.
Virtually every child becomes an expert in natural language, motor control, and social interaction, while learning rich
intuitive theories of kinship, taxonomy and physics. Many of those
children will grow into adults with competence in cooking, calculus, tennis, drawing
pictures, or writing software.
Expertise -- the ability to quickly solve new problems in a domain -- is crucial.
Human learning goes far beyond just memorizing a large set of facts or routines,
but involves higher-order skills,
like learning to learn new concepts or
learning to solve new types of problems more quickly and effectively.
Despite great advances in machine learning,
we are still far from
an AI with these abilities.
The modern AI toolkit has given us machines that learn to play challenging games at superhuman levels,
but cannot quickly transfer to similar games like a person would;
or, which can learn to generate convincing English prose,
but which do not then learn to analyze many different languages,
like a field linguist would.
Here, we present a computational model that takes a step toward the goal of building machines that
grow into domain experts,
looking at a number of problem domains that humans can learn and indeed become experts in.

we draw inspiration from cognitive science.
this inspiration says that there are two things.
we do both these things.
hooray for us.
Our model is structured
around the hypothesis that
domain expertise
consists primarily of two ingredients.
First, domain experts have 
explicit, declarative concepts that are powerful yet finely-tuned.
A visual artist has concepts like arcs, spirals, symmetries, and perspectives;
a physicist has concepts like dot products, vector fields, and conservation laws;
and an architect has concepts like arches, supports, and bridges.
Second, experts have implicit, procedural skill in deploying those concepts
to quickly solve new problems:
at a glance, human domain
experts can intuit which compositions of concepts are likely to solve
the task at hand, even before searching for a solution.
Studies of human expertise
find that,
compared to novices,
experts can, at a glance, categorize and classify problems based on the ``deep structure''
of the problem's solution~\cite{chi1981expertise,chi2012seeing}.
In short, we take the stance that expertise means both
having the right explicit concepts, and being able to quickly see how those concepts can be composed into a solution.

Because solutions to many kinds of problems can often be described as
some kind of
program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
our model approaches a specific problem by searching for a program
solving it. It gradually grows its domain-specific knowledge for a
class of problems by assembling a library of code containing concepts
useful for the domain, while jointly training a neural network to
quickly infer, for a specific problem, what code in its library is
likely to solve it. We think of these two learned components as
analogous to the explicit concepts and implicit procedural skill that
human experts develop.  The model, called \systemEnding, is an
instance of what in the machine learning community is called a
wake-sleep algorithm, where for us, `waking' corresponds to solving
problems, while a pair of `sleep' cycles correspond to improving the
library of explicit concepts and training the neural net to improve
implicit procedural knowledge.

@@

\system works through a novel kind
of wake/sleep or `dream' learning~\cite{hinton1995wake}, iterating
through a wake cycle -- where it solves \textbf{tasks} by writing programs (Figure~\ref{threeCycles} top)
-- and a pair of sleep cycles, both of which are loosely biologically
inspired by the different memory consolidation processes that occur during sleep~\cite{rasch2013sleep}. The first sleep cycle, which we refer to as
\textbf{abstraction}, grows the library of code by replying
experiences from waking and consolidating them into new code
abstractions (Figure~\ref{threeCycles} left).  This cycle is loosely inspired by the formation of declarative abstractions
during slow-wave sleep memory consolidation~\cite{DUDAI201520}.  The second
sleep cycle, which we refer to as \textbf{dreaming}, improves the
agents procedural knowledge of how to write code by training a neural network to
help quickly search for programs. The neural net is trained on replayed
experiences as well as `fantasies', or sampled programs, built from the learned
library (Figure~\ref{threeCycles} right).  These two kinds of dreams are inspired by the distinct
episodic replay and hallucination components of dream
sleep~\cite{fosse2003dreaming}.

Viewed as a probabilistic inference problem, \system observes a set of
tasks, written $X$, and infers both a program solving each task, as
well as a prior distribution over programs likely to solve tasks in
the domain (Figure~\ref{threeCycles} middle). This prior is encoded by
a library, written $\mathcal{D}$, which combined with a
learned weight vector $\theta$
defines a generative model over programs
(Appendix~\ref{sampleProgram}).  The neural network learns to invert
this generative model by predicting, conditioned on a task, a
``posterior'' distribution over programs likely to solve that specific
task,
thus functioning as a \textbf{recognition model} and trained jointly with the generative model, in the spirit of the Helmholtz machine~\cite{hinton1995wake}.
Writing $Q(p|x)$ for the approximate
posterior predicted by the recognition model, wake/sleep cycles correspond to iteratively (and
approximately) solving for
\begin{align}
  p_x &= \argmax_{\substack{p\\Q(p|x)\text{ is large}}} \probability[x|p]\probability[p|\mathcal{D},\theta]\text{, for each task $x\in X$}&&\text{\emph{Wake}}\\
  \mathcal{D},\theta &= \argmax_{\mathcal{D},\theta}\probability\left[\mathcal{D},\theta \right]\prod_{x\in X}\sum_{p\text{ found during waking}} \probability[x|p]\probability[p|\mathcal{D},\theta]&&\text{\emph{Sleep: Abstraction}}\\
  Q(p|x)&\approx \probability[p|x,\mathcal{D},\theta]&&\text{\emph{Sleep: Dreaming}}  
\end{align}
\noindent which serves to maximize a lower bound on the posterior over $(\mathcal{D},\theta)$ given $X$ (Appendix~\ref{probabilisticAppendix}).

@@

    \\\bottomrule \\\\
    \multicolumn{2}{c}{
      \begin{tabular}{ccc}
        Dreams before learning&Dreams after learning (random)&Dreams after learning (selected)\\
          \fbox{\includegraphics[width = 0.3\textwidth]{figures/dreams/beforeLearning25}}&
          \fbox{\includegraphics[width = 0.3\textwidth]{figures/dreams/final25}}&
          \fbox{\includegraphics[width = 0.3\textwidth]{figures/dreams/cherry_picked/montage}}
      \end{tabular}
    }
  \end{tabular}
  \caption{\textbf{Top}: 48 (out of 160) LOGO graphics tasks. Agent writes a program controlling a `pen' that draws the target picture. \textbf{Middle}: Example learned library routines. Agent learns parametric routines for drawing families of curves (left) as well as subroutines that take entire programs as input (right). Each row of images on the left is the same code executed with different parameters. Each image on the right is the same code executed with different parameters and with a different subprogram provided as input. \textbf{Bottom}: 25 dreams, or sampled programs, from library before (left column) and after (middle \& right) learning. Blue: where the agent started drawing. Pink: where the agent ended drawing.}\label{logoMaster}
  \end{figure}

@@

  \end{tabular}
  \setlength{\tabcolsep}{6pt} % default is 6pt
  \caption{\textbf{Left:} Learning a language for physical laws starting with recursive list routines like \code{map} and \code{fold}. Agent observes input/outputs of latent functions encoding physical laws (top left, 20 input/outputs per task). Vectors are represented as lists of real numbers. Physical constants expressed in Planck units. \textbf{Right:} Learning a language for recursive list routines starting with only recursion and primitives found in 1959 Lisp. Agent rediscovers \code{map}, \code{fold}, \code{unfold}, etc. For full tasks and learned libraries see Appendix~\ref{physicsAppendix,appendixMcCarthy}}\label{librariesToLanguages}
  \end{figure}

@@

Our model draws on ideas first elucidated by the
Exploration-Compression family of algorithms~\cite{Dechter:2013:BLV:2540128.2540316,ecc,lazaro2019beyond,DBLP:conf/icml/LiangJK10,solomonoff1989system,schmidhuber2004optimal,ozkural2011teraflop},
which alternate
between searching, or `exploring', a space of solutions, and then
updating their search procedure by compressing solutions found during
exploration.
EC : not only one difference. we actually incorporate several improvements.
for example our compression is very sophisticated.
it is a fancy kind of re- factoring.
exploring is analogous to waking except that we also are better here because we can break symmetries and have a deep network.
Both of our sleep phases can be interpreted as a kind of
compression: abstraction aims to compactly refactor code, while the
recognition model aims to encode a program in as few bits as possible,
conditioned on a task. What \system contributes here is the joint capture of
declarative aspects of domain knowledge -- considered in prior EC works -- and
the equally crucial procedural knowledge,
both of which human experts seem to need.
also it turns out that computers need these two too!!!

@@

Algorithm~\ref{theWholeThing} specifies how we integrate wake and
sleep cycles.
\begin{algorithm}
  \caption{Full \system algorithm}\label{theWholeThing}
  \begin{algorithmic}[1]
    \State \textbf{function} \systemEnding$(\mathcal{D}, X)$:
    \State {\bfseries Input:} Initial DSL $\mathcal{D}$, tasks $X$
    \State \textbf{Output:} Infinite stream of DSLs, recognition models, and beams
    \State \textbf{Hyperparameters:} Batch size $B$, enumeration timeout $T$, maximum beam size $F$
    \State $\theta\gets $uniform distribution
    \State $\mathcal{B}_x\gets\varnothing $, $\forall x\in X$\Comment{Initialize beams to be empty}
    \While{true}\Comment{Loop over epochs}
    \State shuffle $\gets$ random permutation of $X$\Comment{Randomize minibatches}
    \While{shuffle is not empty}\Comment{Loop over minibatches}
    \State batch $\gets$ first $B$ elements of shuffle\Comment{Next minibatch of tasks}
    \State shuffle $\gets$ shuffle with first $B$ elements removed
    \State $\forall x\in \text{batch}$: $\mathcal{B}_x\gets\mathcal{B}_x\cup \left\{p\;|\;p\in \text{enumerate}(\probability[\cdot |\mathcal{D},\theta],T)\text{ if }\probability[x|p] > 0 \right\}$\Comment{Wake}
    \State Train $Q(\cdot |\cdot )$ to minimize $\mathcal{L}^{\text{MAP}}$ across all $\left\{\mathcal{B}_x \right\}_{x\in X}$\Comment{Dream Sleep}
    \State $\forall x\in \text{batch}$: $\mathcal{B}_x\gets\mathcal{B}_x\cup \left\{p\;|\;p\in \text{enumerate}(Q(\cdot |x),T)\text{ if }\probability[x|p] > 0 \right\}$\Comment{Wake}
    \State $\forall x\in \text{batch}$: $\mathcal{B}_x\gets\text{ top }F\text{ elements of }\mathcal{B}_x\text{ as measured by }\probability[\cdot |x,\mathcal{D},\theta]$\Comment{Keep top $F$ programs}
    \State $\mathcal{D},\theta,\left\{\mathcal{B}_x \right\}_{x\in X}\gets $\textsc{Consolidate}$(\mathcal{D},\theta,\left\{\mathcal{B}_x \right\}_{x\in X})$\Comment{Consolidation Sleep}
    \State \textbf{yield} $(\mathcal{D},\theta)$, $Q$, $\left\{\mathcal{B}_x \right\}_{x\in X}$ \Comment{Yield the updated DSL, recognition model, and solutions found to tasks}
    \EndWhile
    \EndWhile

@@

\begin{algorithm}%[H]
  %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
  \caption{Library Induction Algorithm}
  \label{grammarInductionAlgorithm}
  \begin{algorithmic}[1]
    \State {\bfseries Input:} Set of beams $\{\mathcal{B}_x\}$
    %         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
    \State \textbf{Output:} Library $\mathcal{D}$, weight vector $\theta$
    \State $\mathcal{D}\gets$ every primitive in $\{\mathcal{B}_x\}$
    \While{true}         
    \State $\forall p\in \bigcup_{x}\mathcal{B}_x: $ $v_p\gets I\beta(p)$ \Comment{Construct a version space for each program}
    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{B}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(v_p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood under $(\mathcal{D}',\theta)$}
    \State Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
    \State Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*(\mathcal{D})) - \|\theta^*(\mathcal{D})\|_0$ \Comment{objective function}
    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{B}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new components}
    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new libraries}
    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new library}
    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to library led to a better score}
    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better library. Update library.}
    \State $\forall x\;:\;\mathcal{B}_x\gets\left\{\text{\textsc{refactor}}(v_p|\mathcal{D})\;:\; p\in \mathcal{B}_x\right\}$\Comment{Refactor beams in terms of new library}
    \EndWhile
  \end{algorithmic}
\end{algorithm}

 \subsubsection{Refactoring code with version spaces}\label{appendixVersion}
 Formally, a version space is either:
\begin{itemize}
\item A deBuijn\footnote{deBuijn indices are an alternative way of naming variables in $\lambda$-calculus. When using deBuijn indices, $\lambda$-abstractions are written \emph{without} a variable name, and variables are written as the count of the number of $\lambda$-abstractions up in the syntax tree the variable is bound to. For example, $\lambda x.\lambda y. (x\;y)$ is written $\lambda\lambda (\$1\;\$0)$ using deBuijn indices. See~\cite{pierce} for more details.} index: written $\$i$, where $i$ is a natural number
\item An abstraction: written $\lambda v$, where $v$ is a version space
\item An  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
\item A  union: $\uplus V$, where $V$ is a set of version spaces
\item The empty set, $\varnothing$
\item The set of all $\lambda$-calculus expressions, $\Lambda$
\end{itemize}
The purpose of a version space to compactly represent a set of programs.
We refer to this set as the \textbf{extension} of the version space:
\begin{definition}
  The \textbf{extension} of a version space $v$ is written $\denotation{v}$
  and is defined recursively as:
  \begin{align*}
    \denotation{\$i}& = \left\{\$i \right\}&
    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}&
    \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}&
    \denotation{\varnothing}& = \varnothing&
    \denotation{\Lambda}& = \Lambda
    \end{align*}
\end{definition}
Version spaces also support efficient membership checking, which we
write as  $e\in \denotation{v}$.
Important for our purposes, it is also efficient to refactor the
members of a version space's extension in terms of a new library.
We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
\begin{align*}
  \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
    e\text{, if $e\in \mathcal{D}$ and $e\in \denotation{v}$. Exploits the fact that $\denotation{e}\in v$ can be efficiently computed.}\\
    \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
  \end{cases}
\end{align*}\vspace{-\baselineskip}%
\begin{align*}
  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}&%\\\nonumber
  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\\nonumber
  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})&%\\\nonumber
  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\text{size}(e|\mathcal{D})\nonumber
  \end{align*}
where $\text{size}(e|\mathcal{D})$ for program $e$ and library $\mathcal{D}$
is the size of the syntax tree of $e$,
when members of $\mathcal{D}$ are counted as having size 1.
Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\text{size}(p|\mathcal{D})$.

@@

\subsubsection{Computational complexity of library learning}

How long does each update to the library in
Algorithm~\ref{grammarInductionAlgorithm} take?  Constructing the
version spaces takes time linear in the number of programs (written
$P$) in the beams (Algorithm~\ref{grammarInductionAlgorithm}, line
5), and, in the worst case, exponential time as a function of the
number of refactoring steps $n$ --- but we bound the number of steps
to be a small number (typically $n = 3$).  Writing $V$ for the number
of version spaces, this means that $V$ is $O(P2^n)$.  The number of
proposals (line 10) is linear in the number of distinct version
spaces, so is $O(V)$.  For each proposal we have to refactor every
program (line 6), so this means we spend $O(V^2) = O(P^22^n)$ per library
update.  In practice this quadratic dependence on $P$ (the number of
programs) is prohibitively slow.  We now describe a linear time
approximation to the refactor step in
Algorithm~\ref{grammarInductionAlgorithm} based on beam search.
